### 2019-01-16 Datelife_paper1.Rmd
last paragraph of intro

<!--INTRODUCTION-->
Divergence time of lineages<!--named biological entities, taxon ages?--> constitutes in many ways the main knowledge necessary for evolutionary understanding.
<!--Along with phylogenetic relationships, it is the most important piece of information allowing to test alternative evolutionary hypothesis in many areas of research, from developmental to ecological questions.
Along with migration rates and patterns, it is fundamental to test biogeographical hypothesis-->
Coupled to species number and distribution, it is the basic information for the study of diversification processes, such as the tempo and mode of speciation and extinction, crucial for the understanding of how biodiversity patterns are shaped across space and time [@Morlon2014].

When organisms are preserved in a fossil form, a time frame of taxon origin can be obtained directly from the age of rock strata.
However, not all organisms fossilize well or at all.
Fossilization success alone is highly circumstancial, and varies depending on a number of parameters including the nature of the habitat, population size, species range breadth and physical characteristics of the organism. <!--fossil recovery is also highly circumstancial-->
Thus, relying only on the fossil record to obtain a time frame of lineage divergence<!--taxon ages--> for all life is<!--, alas,--> not possible.
Relative rate of DNA or aminoacid substitution constitutes another important source of information on lineage divergence. <!--describe the two components of rel subst rate: which, contain tne information on absolute rate of substitution and absolute time.-->
It is usually obtained from hypothesis of character homology (alignments) when reconstructing phylogenetic relationships. Molecular dating techniques use external data such as absolute time calibrations (e.g., fossils, geologic events) or absolute substitution rates <!--birds--> to generate dated phylogenies (chronograms) which contain information on absolute times of node divergence and taxon ages. <!--With molecular dating methods, an external calibration point is still necessary to estimate the absolute time of origin. Hence, the accumulation of data on fossils for calibration is also continual-->
<!--This method has been particularly useful to obtain time frame of origin of organisms with poor or no fossil record, but is also largely used for organisms with rich fossil records-->
<!--Models allow to take into account the nature of the fossil record: Given the characteristics of the fossilization process, the quality of dates inferred directly from it is not always considered the best.-->

In the past decades, the possibility to obtain DNA sequences in large quantities from a wide variety of organisms became a reality, which, coupled to methodological development in phylogenetic and dating inference, allowed the application of molecular dating methods on a very large amount and diversity of organisms, greatly increasing the quantity of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of both fossil and molecular-based data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees)
names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, which coupled to better sharing data practices, are greatly contributing to the accumulation in number and type of available data on taxon ages.

<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->
One inspiration for the DateLife project was the TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017]. This project has aggregated chronograms from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow. However, even in this gold standard resource, the included taxa only include between 0.097 and 3.236% of total species diversity (following taxonomic expert opinion on the global, extant species numbers, which ranges from 3 to 100 million species [Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from across the tree of life, versus more specialized chronograms focusing on plants [PHYLOMATIC], birds [JETZ ET AL BIRDTREE.ORG], and other groups. Users can choose between a web interface or a mobile app to receive information on divergence times for the evolutionary history of a lineage, pairs of taxa, all lineages within a taxon, or a list of taxa. As a science communication tool, TimeTree project is a very powerful tool: it has a friendly graphical interface, with informative while colorful outputs, that allows the general public to satisfy curiosity regarding their favorite organism or group of them. It is of limited utility for scientific studies, however. The thousands of trees that have been entered are unavailable for examination or reuse; according to their FAQ, methods for allowing download of the data have been under discussion since 2015. Also, there is no Application Programming Interface (API) allowing programmatic access to any data. Moreover, large-scale, automated, data-mining is not allowed under the site's terms of use. The nearly hundred thousand taxon chronogram is not available with [@Kumar2017] or the TimeTree website, though the still substantial chronogram from [@Hedges2015] is available at OToL.

With this inspiration, a prototype DateLife service was developed over a series of phylotastic hackathons [__CITE__] at the National Evolutionary Synthesis Center. NSF funding allowed for the further development leading to this paper. A core goal has been openness of both the data sources and the code underlying the analyses.

@@

Total ssp number estimated to be between 3 and 100 million species, as mentiones in [@Mora2011]:
ttol_spp <- 97085
(ttol_spp * 100) / 3e+6  # 3.236167
(ttol_spp * 100) / 100e+6  # 0.097085
-->
<!--
While its node-time search is very powerful, using available data to infer divergence time from lineages that have never been sequenced before, its timetree option is less powerful, their main flaw being that it does not provide source data.
However, their primary database is not fully open, giving information on a per node basis. So, it is not possible to easily access information from original studies for a group of taxa (One would need to perform a pair of taxa search for all pairs from the group of interest).
Divergence time estimates for all taxa in the TToL are precomputed and stored in the TimeTree database
All searches allow species names and higher group names. Common names are allowed in all searches but a list of taxa.
one-taxon search was nor working for dog, though.
1) ttol one lineage: it gives the evolutionary timeline of a lineage, showing the age of main divergence events occuring along the lineage evolutionary history, i.e., when it shared a comon ancestor with all main lineages from which it diverged a long time ago.
It shows results only graphically on the website.
2) ttol parwise search: allows common names, species names and higher group names
    how it handles names missing from tree: it looks in the NCBI taxonomy for the closest relative of the species or higher taxa. Then uses the closest relative to get the lineage pair age from ttol.
3) ttol group of taxa
4) ttol list of taxa: Cannot get it to work; does not work with cat, dog, turtle search, felidae, hominidae, phrynops search, some_ants species name search. What kind of files does it accepts: txt, rtf, doc, docx?
-->

@@

Despite its importance, analytical tools to summarize available information on taxon ages for the scientific community are still lacking.
Data might not have been exploited because:
Data is in different repositories and formats;
Lineage names are different among studies and difficult to reconcile;
Taxonomy is also different among studies and difficult to reconcile.
<!--Data curation is necessary at some point (is this always true?)-->
Also, data curation is an important part of any biological study. The research community considers it as an important or even crucial step before data analysis. Hence, automated processes for large data analysis are frequently received with skepticism.
Recent work on this area (e.g., supersmart<!--and, which others?-->) aims to:
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of data curation is still necessary.
<!--General issues with dating techniques:-->
Choosing correct fossils for calibration requires a lot of expertise and knowledge on the group.
Incorrect use of fossils can generate severe bias in dating results [@Sauquet2012c].
DateLife palliates this by using only information available from already published studies, which are ideally constructed using robust information, such as sequence data and curated fossil calibrations.
<!--Data curation is largely based on taxonomic knowledge; or at least on the interplay between taxonomy and phylogenetics-->
<!--Importance of datelife:-->
DateLife can summarize this information in several formats that can be easily inspected by users.
This allows rapidly obtaining a time frame of lineage divergence for a wide number of taxa.
DateLife can also generate chronograms for taxa with little available information, by using the available data as calibration points.
DateLife is the main service for scaling phylogenetic trees in Phylotastic! system [@Stoltzfus2013] <!--to construct phylogenetic trees on the fly [http://phylotastic.org/].-->
It can be used through an R package <link to documentation>, a web interface (<http://www.datelife.org/query/>) and an API<!--(still not up, right?)-->.

@@

<!--(BOLD and OToL species names are homogeneous?-->

DateLife is a service for searching and processing information on ages of any taxon <!--or group of taxa-->of interest, available in chronograms from public data repositories of published peer reviewed studies. With this data, it can also generate new taxon age information<!--by using expert available phylogenies and scaling them to absolute time using information on ages obtained from published chronograms,--> by calling a variety of external services and proved methods.
It only requires as input a set of taxon names, in the form of a comma separated listing or vector, or of a phylogeny with taxon names on the tips.
Taxon names can correspond to binomial species names or clades<!--common names (will we implement it?)-->.
When taxon names are clades, DateLife pulls all accepted species names within the clade (up to OToL's limit of _____ species) from OToL's reference taxonomy using a service of rphylotastic R package<!--citation-->.
Names belonging to subspecies or any other infraspecific category are treated as species.
DateLife can process input names with the taxon name resolution service (TNRS), <!--which "corrects misspelled names and authorities, standardizes variant spellings, and converts nomenclatural synonyms to accepted names" (Boyle et al. 2013)-->which corrects misspelled names or typos, and standardizes variation in spelling and synonyms [@Boyle2013], increasing the probability to correctly find the queried taxa in the chronogram database. DateLife uses TNRS to compare names against OToL's reference taxonomy using a service from the R package rotl [@Michonneau2016]. <!--we'll implement the option to use other reference taxonomies in the future -->

DateLife's main function searches taxon names across the chronogram database specified by the user. At the moment, it queries chronograms from the OToL [@Hinchliff2015]<!-- there is an earlier reference from 1994 but it does not exist in its journal database: Sanderson MJ, Donoghue MJ, Piel WH, Eriksson T: TreeBASE: a prototype database of phylogenetic analyses and an interactive tool for browsing the phylogeny of life. Am J Bot. 1994, 81 (6): 183--> repository.
DateLife identifies chronograms having at least two taxon names, and subset them to contain only the taxa of interest. It then stores taxon age information from each chronogram individually as a patristic matrix, named with the citation of the original study. This format allows a rapid summary in a number of different ways, including: 1) citations of the original studies containing the subset chronograms, 2) a list of subset chronograms' mrca ages, 3) a list of subset chronograms in newick or phylo format, 4) a table containing all information in html or R's data frame format, or 5) as a single chronogram summarized from subset chronograms using the supertree Super Distance Matrix (SDM) approach [@Criscuolo2006] or using the median of branch lengths. <!--(how to better explain the latter?? Also, I just discovered the median gives non ultrametric trees...).-->

@@

Problems and caveats:

@@

DateLife is free and open source and it can be used through its current website <http://www.datelife.org/query/>, or through Phylotastic!'s web portal <http://phylo.cs.nmsu.edu:3000/>.
DateLife can also be used locally through its R package. The stable version is available for installation from the CRAN repository (<https://cran.r-project.org/package=datelife>) using the command `install.packages(pkgs = "datelife")` from R. Development versions are available from GitHub repository (<https://github.com/phylotastic/datelife>) and can be installed using the devtools R package command `install_github("phylotastic/datelife")`.

@@

<!--**Acknowledgements**-->
<span style="font-variant:small-caps;">
Acknowledgements
</span>

>>>>

<!--INTRODUCTION-->
Time of lineage divergence<!--named biological entities, taxon ages?--> constitutes in many ways the fundamental/main knowledge necessary for evolutionary understanding.
<!--Along with phylogenetic relationships, it is the most important piece of information allowing to test alternative evolutionary hypothesis in many areas of research, from developmental to ecological questions.
Along with migration rates and patterns, it is fundamental to test biogeographical hypothesis-->
Coupled to species number and distribution, it is the main information necessary for the study of diversification processes (i.e., the tempo and mode of speciation and extinction),
<!-- crucial/vital/critical/essential  -->central for the understanding of how biodiversity patterns are shaped across space, time and clades [@Morlon2014].
Evolutionary understanding also relies on comparative studies, for which knowing the time context for all life is crucial. Efforts to have a whole tree of life have been great and here are some examples.
<!--INTRODUCTION: 2nd paragraph-->
When organisms are preserved in a fossil form, a time frame of taxon origin can be obtained directly from the age of rock strata.
However, not all organisms fossilize well or at all.
Fossilization success alone is highly circumstancial, and varies depending on a number of parameters including the nature of the habitat, population size, species range breadth and physical characteristics of the organism, all of which greatly vary across different organisms and through time. <!--fossil recovery is also highly circumstancial-->
Thus, relying only on the fossil record to obtain a time frame of lineage divergence<!--taxon ages--> for all life is<!--, alas,--> not possible.

Another caveat is that we usually cannot have a point age estimates of lineage origin because of the nature of the fossilization process: when a fossil first appears on the fossil record it is not necessarily because the lineage just originated, it might just be bc the conditions for fossilization started at that point but the organism might have been around for a long time before that. Because of this, we use ranges from the age of strata where fossils are found. But this does not solve the problem. The organism could even have been around previously, but not fossilize under the conditions prevailing around the time of formation of rock strata below its first appearance in the fossil record. In this sense, fossil ages can only be considered as minimum time of origin of lineages.
<!--D Bapst work?-->

Another source that has been widely used to inform about the timing of lineage origin is the relative rate of DNA or aminoacid substitution.
It is estimated from hypothesis of character homology (alignments) for reconstructing phylogenetic relationships.
<!--describe the two components of rel subst rate? bc they contain the information on absolute rate of substitution and absolute time.-->
Molecular dating techniques use external data such as absolute time calibrations (e.g., fossils, geologic events) or absolute substitution rates <!--obtained in the lab?? birds-->
to <!--dissociate rel subs rates components and--> generate dated phylogenies (chronograms) which contain information on absolute times of node divergence and taxon ages. <!--With molecular dating methods, an external calibration point is still necessary to estimate the absolute time of origin. Hence, the accumulation of data on fossils for calibration is also continual-->
<!--This method has been particularly useful to obtain time frame of origin of organisms with poor or no fossil record, but is also largely used for organisms with rich fossil records-->
<!--Models allow to take into account the nature of the fossil record: Given the characteristics of the fossilization process, the quality of dates inferred directly from it is not always considered the best.-->
<!--INTRODUCTION: 3rd paragraph-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed the application of molecular dating methods on a very large amount and diversity of organisms, greatly increasing the quantity of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of both fossil and molecular-based data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE and Open Tree of Life (OToL).
OToL alone holds more than 200 chronograms. <!--cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees)
names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
Methods to include living and fossil lineages are in continued development and increased usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, which coupled to better sharing data practices, are greatly contributing to the accumulation in number and type of available data on taxon ages.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->

<!--INTRODUCTION: 4th paragraph-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow. However, even in this gold standard resource, the included taxa only encompass between 0.097 and 3.236% of total species diversity (following taxonomic expert opinion on the global, extant species numbers, which ranges from 3 to 100 million species [Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from across the tree of life, versus more specialized chronograms focusing on plants [PHYLOMATIC], birds [JETZ ET AL BIRDTREE.ORG], and other groups. Users can choose between a web interface or a mobile app to receive information on divergence times for the evolutionary history of a lineage, pairs of taxa, all lineages within a taxon, or a list of taxa. As a science communication tool, TimeTree project is very powerful: it has a friendly graphical interface, with informative and colorful outputs, that allows the general public to satisfy curiosity regarding a particular organism of interest or group of them. It is of limited utility for scientific studies, however. The thousands of trees that have been entered are unavailable for examination or reuse; according to the creators (see TimeTree web FAQ), methods for allowing data downloading have been under discussion for the past several years yet the primary data remain closed. Moreover, there is no Application Programming Interface (API) allowing programmatic access to any data, greatly impairing the possibility of large-scale, automated data-mining, which is not allowed under TimeTree website's terms of use. The nearly hundred thousand taxon summary chronogram generated from TimeTree resources is not available with its publication [@Kumar2017] or the TimeTree website, though the still substantial chronogram from a previous publication [@Hedges2015] was made available at OToL.

With this inspiration, a prototype DateLife service was developed over a series of phylotastic hackathons [__CITE__] at the National Evolutionary Synthesis Center. NSF funding allowed for further development leading to this paper. A core goal of DateLife is openness of both the data sources and the code underlying the analyses, so scientific community can take advantage of this tool to leverage available information to the advancement of discoveries in biology.

@@

Total ssp number estimated to be between 3 and 100 million species, as mentiones in [@Mora2011]:
ttol_spp <- 97085
(ttol_spp * 100) / 3e+6  # 3.236167
(ttol_spp * 100) / 100e+6  # 0.097085
-->
<!--
While its node-time search is very powerful, using available data to infer divergence time from lineages that have never been sequenced before, its timetree option is less powerful, their main flaw being that it does not provide source data.
However, their primary database is not fully open, giving information on a per node basis. So, it is not possible to easily access information from original studies for a group of taxa (One would need to perform a pair of taxa search for all pairs from the group of interest).
Divergence time estimates for all taxa in the TToL are precomputed and stored in the TimeTree database
All searches allow species names and higher group names. Common names are allowed in all searches but a list of taxa.
one-taxon search was not working for dog, though.
1) ttol one lineage: it gives the evolutionary timeline of a lineage, showing the age of main divergence events occuring along the lineage evolutionary history, i.e., when it shared a comon ancestor with all main lineages from which it diverged a long time ago.
It shows results only graphically on the website.
2) ttol parwise search: allows common names, species names and higher group names
    how it handles names missing from tree: it looks in the NCBI taxonomy for the closest relative of the species or higher taxa. Then uses the closest relative to get the lineage pair age from ttol.
3) ttol group of taxa
4) ttol list of taxa: Cannot get it to work; does not work with cat, dog, turtle search, felidae, hominidae, phrynops search, some_ants species name search. What kind of files does it accepts: txt, rtf, doc, docx?
-->

@@

Despite its great importance, analytical tools to summarize available information on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of existing data. First, original chronograms available publicly are scattered across various repositories (otol tree store, dryad, treebase, journals supplementary data) usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies and manual curation of that is usually necessary.
Third, data curation
Recent advances on this area (e.g., supersmart<!--and, which others?-->) aim to:
Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;
Problems or downsides: This might be time consuming for large groups and a lot of data curation and knowledge on the group of interest is still necessary. For example, <!--General issues with dating techniques:--> choosing correct fossils for calibration requires a lot of expertise and knowledge on the group. An incorrect use of fossils can generate severe bias in dating results [@Sauquet2012c].
<!--Data curation is necessary at some point (is this always true?)-->
Hence, data curation is still an important part of any biological study. The research community considers it as an important or even crucial step before data analysis. Hence, automated processes for large data analysis are frequently received with skepticism.
<!--INTRODUCTION: 5th paragraph-->

DateLife palliates this by only using information available from already published studies, which are ideally constructed using robust information, such as sequence data and thoughtfully curated fossil calibrations.
<!--Data curation is largely based on taxonomic knowledge; or at least on the interplay between taxonomy and phylogenetics-->
<!--Importance of datelife:-->
DateLife can summarize this information in several formats that can be easily inspected by users.
This allows rapidly obtaining a time frame of lineage divergence for a wide number of taxa.
DateLife can also generate chronograms for taxa with little available information, by using the available data as calibration points.
DateLife is the main service for scaling phylogenetic trees in Phylotastic! system [@Stoltzfus2013] <!--to construct phylogenetic trees on the fly [http://phylotastic.org/].-->
It can be used through an R package <link to documentation>, a web interface (<http://www.datelife.org/query/>) and an API<!--(still not up, right?)-->.
In here we present the first release of DateLife.
It contains an improved databse of chronograms, more methods to summarize trees, and new functions to visualize data.
It also allows comparison of summary methods.

@@

<!--(BOLD and OToL species names, are they homogeneous/standardized?-->

DateLife is a service for searching and processing information on ages of any taxon <!--or group of taxa-->of interest, available in chronograms from public data repositories of published peer reviewed studies. From these data, it can also generate new taxon age information<!--by using expert available phylogenies and scaling them to absolute time using information on ages obtained from published chronograms,--> by calling a variety of external services and tested algorithms.
It takes advatage of the rotl, ape, and geiger packages to gather, process, and present information.

It only requires a set of taxon names as input, in the form of a comma separated listing or vector, or of a phylogeny with taxon names on the tips.
Taxon names can correspond to binomial species names or clades<!--common names (will we implement it?)-->.
When taxon names are clades, DateLife pulls all accepted species names within the clade (up to OToL's limit of _____ species) from OToL's reference taxonomy using a service of rphylotastic R package<!--citation-->.
Names belonging to subspecies or any other infraspecific category are treated as species.
DateLife can process input names with the taxon name resolution service (TNRS), <!--which "corrects misspelled names and authorities, standardizes variant spellings, and converts nomenclatural synonyms to accepted names" (Boyle et al. 2013)-->which corrects misspelled names or typos, and standardizes variation in spelling and synonyms [@Boyle2013], increasing the probability to correctly find the queried taxa in the chronogram database. DateLife uses TNRS to compare names against OToL's reference taxonomy using a service from the R package rotl [@Michonneau2016]. <!--we'll implement the option to use other reference taxonomies in the future -->

DateLife's main function searches taxon names across the chronogram database specified by the user. At the moment, it queries chronograms from the OToL [@Hinchliff2015]<!-- it cannot query chronograms from TreeBASE -there is an earlier reference from 1994 but it does not exist in its journal database: Sanderson MJ, Donoghue MJ, Piel WH, Eriksson T: TreeBASE: a prototype database of phylogenetic analyses and an interactive tool for browsing the phylogeny of life. Am J Bot. 1994, 81 (6): 183--> repository.
DateLife identifies chronograms having at least two taxon names, and subset them to contain only the taxa of interest. It then stores taxon age information from each chronogram individually as a patristic matrix, named with the citation of the original study. This format allows a rapid summary in a number of different ways, including: 1) citations of the original studies containing the subset chronograms, 2) a list of mrca ages of subset chronograms, 3) a list of complete subset chronograms in newick or phylo format, 4) a table containing all information retrieved in html or R's data frame format, or 5) a single chronogram summarized from subset chronograms using the Super Distance Matrix (SDM) supertree construction approach [@Criscuolo2006] or using the median of branch lengths. <!--(how to better explain the latter?? Also, I just discovered the median gives non ultrametric trees...).-->

@@

Problems and caveats:
Not many databases, only OToL
Why TreeBase is not very useful for us? Be precise.

@@

DateLife is free and open source and it can be used through its current website <http://www.datelife.org/query/>, or through Phylotastic!'s web portal <http://phylo.cs.nmsu.edu:3000/>. 
RStudio's Shiny Server and the shiny package open infrastructure are used to maintain the former. Also Docker.
DateLife can also be used locally through its R package. The stable version is available for installation from the CRAN repository (<https://cran.r-project.org/package=datelife>) using the command `install.packages(pkgs = "datelife")` from R. Development versions are available from GitHub repository (<https://github.com/phylotastic/datelife>) and can be installed using the devtools R package command `install_github("phylotastic/datelife")`.

@@

<!--**Acknowledgements**-->
<span style="font-variant:small-caps;">
Acknowledgements
</span>
We thank colleagues (students and postdocs) at the O'Meara Lab at the University of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of trees used for DateLife.
The many scientists who publish their chronograms in an open, reusable form, and the scientists who curate them for deposition in OpenTree.
The US National Science Foundation (NSF) for funding nearly all the above, in addition to the ABI grant that funded this project itself.
